---
title: "Introducing 'brat'"
subtitle: "A wrapper for the rapid annotation tool"
author: "Andreas Blätte (andreas.blaette@uni-due.de)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introducing 'brat'}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---


## Requirements

Apart from the brat package, we load NLP which offers very useful data structure to get the data required by brat.

```{r}
library(brat) 
library(NLP)
```


## Minimal working example

A first minimal example uses the data included in the tutorial ['embedding brat visualisations]('https://brat.nlplab.org/embed.html).

First, we define document data that include the text and spans of text to be annotated.

```{r}
doc_data <- list(
   text = "Ed O'Kelley was the man who shot the man who shot Jesse James.",
   entities = list(
     list('T1', 'Person', list(c(0, 11))),
     list('T2', 'Person', list(c(20, 23))),
     list('T3', 'Person', list(c(37, 40))),
     list('T4', 'Person', list(c(50, 61)))
   )
)
```

Second, we define collection data, i.e. definitions that go beyond the individual documents applicable for an entire collection of documents.

```{r}
coll_data <- list(
  entity_types = list(entity("Person", bgColor = RColorBrewer::brewer.pal(8, "Accent")[1]))
)
```

This is all it takes for a first brat visualisation.

```{r, eval = TRUE}
brat(doc_data = doc_data, coll_data = coll_data, width = "700px")
```


